---
title: "personal task"
author: "m.pryidun"
date: "2022-11-21"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sentiment Analysis 
 Sentiment analysis of five different novels using dataset was published in Saif M. Mohammad and Peter Turney (2013), Crowdsourcing a Word-Emotion Association Lexicon.


Install packages

```{r}
# install packages
# install.packages("dplyr")
#install.packages("stringr")
#install.packages("tidyr")
#install.packages("tibble")
#install.packages("tidytext")
#install.packages("textdata")
#install.packages("Hmisc")
#install.packages("sentimentr")
#install.packages("zoo")
#install.packages("flextable")
#install.packages("remotes")
#remotes::install_github("rlesur/klippy")
```

```{r}
# activate packages
library(dplyr)
library(stringr)
library(tidyr)
library(tibble)
library(tidytext)
library(textdata)
library(Hmisc)
library(sentimentr)
library(zoo)
library(flextable)
# activate klippy for copy-to-clipboard button
klippy::klippy()
```

## Read data

Load five pieces of literature:

```{r pressure, echo=FALSE}
darwin <- base::readRDS(url("https://slcladal.github.io/data/origindarwin.rda", "rb"))
twain <- base::readRDS(url("https://slcladal.github.io/data/twainhuckfinn.rda", "rb"))
orwell <- base::readRDS(url("https://slcladal.github.io/data/orwell.rda", "rb"))
lovecraft <- base::readRDS(url("https://slcladal.github.io/data/lovecraftcolor.rda", "rb"))
```

## Clean data

Write function to clean data

```{r}
txtclean <- function(x, title){
  require(dplyr)
  require(stringr)
  require(tibble)
  x <- x %>%
    iconv(to = "UTF-8") %>%
    base::tolower() %>%
    paste0(collapse = " ") %>%
    stringr::str_squish()%>%
    stringr::str_split(" ") %>%
    unlist() %>%
    tibble::tibble() %>%
    dplyr::select(word = 1, everything()) %>%
    dplyr::mutate(novel = title) %>%
    dplyr::anti_join(stop_words) %>%
    dplyr::mutate(word = str_remove_all(word, "\\W")) %>%
    dplyr::filter(word != "")
}
```

Process and clean texts.
```{r}
darwin_clean <- txtclean(darwin, "darwin")
lovecraft_clean <- txtclean(lovecraft, "lovecraft")
orwell_clean <- txtclean(orwell, "orwell")
twain_clean <- txtclean(twain, "twain")
```
## Data analyzing

Combine the data with the Word-Emotion Association Lexicon

```{r}
novels_anno <- rbind(darwin_clean, twain_clean, orwell_clean, lovecraft_clean) %>%
  dplyr::group_by(novel) %>%
  dplyr::mutate(words = n()) %>%
  dplyr::left_join(tidytext::get_sentiments("nrc")) %>%
  dplyr::mutate(novel = factor(novel),
         sentiment = factor(sentiment))
head(novels_anno,10)
```
Summarize the results of the SA and calculate the percentages of the prevalence of emotions across the books

```{r}
novels <- novels_anno %>%
  dplyr::group_by(novel) %>%
  dplyr::group_by(novel, sentiment) %>%
  dplyr::summarise(sentiment = unique(sentiment),
                   sentiment_freq = n(),
                   words = unique(words)) %>%
  dplyr::filter(is.na(sentiment) == F) %>%
  dplyr::mutate(percentage = round(sentiment_freq/words*100, 1))
head(novels, 10)
```
Visualize the results and show the scores fro each core emotion by book.
```{r}
novels %>%
  dplyr::filter(sentiment != "positive",
         sentiment != "negative") %>%
  ggplot(aes(sentiment, percentage, fill = novel)) +    
  geom_bar(stat="identity",   
           position=position_dodge()) + 
  scale_fill_manual(name = "", values=c("orange", "gray70", "red", "grey30")) +
  theme_bw() +
  theme(legend.position = "top")
```

