---
title: "lab1"
author: "m.pryidun"
date: "2022-11-10"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Task 1. Exploraring a dataset
```{r}
airquality <- read.csv("D:/KNU/DataProcSoft/Data-processing-software/lab1/airquality.csv")
View(airquality)
```
#Question 1. What are the column names of the data frame?
```{r}
cols <- colnames(airquality)    
cols
```
#Question 2. What are the row names of the data frame?
```{r}
rows <- rownames(airquality)
rows
```
#Question 3. Extract the first 6 rows of the data frame and print them to the console
```{r}
head6 <- head(airquality,n=6)
print(head6)
```
#Question 4. How many observations (i.e. rows) are in this data frame?
```{r}
nrow(airquality)
```
#Question 5. Extract the last 6 rows of the data frame and print them to the console
```{r}
tail6 <- tail(airquality,n=6)
print(tail6)
```

#Question 6. How many missing values are in the “Ozone” column of this data frame?
```{r}
which(is.na(airquality$Ozone))
```

#Question 7. What is the mean of the “Ozone” column in this dataset? Exclude missing values (coded as NA) from this calculation.
```{r}
mean ( airquality$Ozone, na.rm  =  TRUE )
```

#Question 8. Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90.
```{r}
identify_rows <- airquality$Ozone >  31  & airquality$Ozone <  90
identify_rows
```

#Question 9. Use a for loop to create a vector of length 6 containing the mean of each column in the data frame (excluding all missing values).
```{r}
means<-c()
for(i in 1:ncol(airquality)) {
    means<-append(means, mean ( airquality[,i], na.rm  =  TRUE ))
}
means
```

#Question 10. Use the apply function to calculate the standard deviation of each column in the data frame (excluding all missing values).
```{r}
apply(airquality, 2, sd)
```
#Question 11. Calculate the mean of “Ozone” for each Month in the data frame and create a vector containing the monthly means (exclude all missing values).
```{r}
tapply(airquality$Ozone, airquality$Month, mean, na.rm = TRUE)
```

#Question 12. Draw a random sample of 5 rows from the data frame
```{r}
set.seed(1)
airquality[sample(nrow(airquality), 5), ]
```


