---
title: "lab4"
author: "m.pryidun"
date: "2022-11-16"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 4. Getting and Clearning Data
The data used in the task was collected from the accelerometers from the Samsung Galaxy S
smartphone. A full description is available at the site where the data was obtained:
http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones
Files for the project you can find in the data folder in getdata_projectfiles_UCI HAR
Dataset.zip zip-archive.
You goal is to prepare tidy data that can be used for later analysis.
```{r}
featureNames <- read.table("UCI HAR Dataset/features.txt")
activityLabels <- read.table("UCI HAR Dataset/activity_labels.txt", header = FALSE)
subjectTrain <- read.table("UCI HAR Dataset/train/subject_train.txt", header = FALSE)
activityTrain <- read.table("UCI HAR Dataset/train/y_train.txt", header = FALSE)
featuresTrain <- read.table("UCI HAR Dataset/train/X_train.txt", header = FALSE)
subjectTest <- read.table("UCI HAR Dataset/test/subject_test.txt", header = FALSE)
activityTest <- read.table("UCI HAR Dataset/test/y_test.txt", header = FALSE)
featuresTest <- read.table("UCI HAR Dataset/test/X_test.txt", header = FALSE)
```

1. Merges the training and the test sets to create one data set.
```{r}
subject <- rbind(subjectTrain, subjectTest)
activity <- rbind(activityTrain, activityTest)
features <- rbind(featuresTrain, featuresTest)
colnames(features) <- t(featureNames[2])
colnames(activity) <- "Activity"
colnames(subject) <- "Subject"
completeData <- cbind(features,activity,subject)
completeData
```


2. Extracts only the measurements on the mean and standard deviation for each measure-ment.
```{r}
columnsWithMeanSTD <- grep(".*Mean.*|.*Std.*", names(completeData), ignore.case=TRUE)
requiredColumns <- c(columnsWithMeanSTD, 562, 563)
dim(completeData)
extractedData <- completeData[,requiredColumns]
dim(extractedData)
```

3. Uses descriptive activity names to name the activities in the data set
```{r}
extractedData$Activity <- as.character(extractedData$Activity)
for (i in 1:6){
extractedData$Activity[extractedData$Activity == i] <- as.character(activityLabels[i,2])
}
extractedData$Activity <- as.factor(extractedData$Activity)
extractedData$Activity
```

4. Appropriately labels the data set with descriptive variable names.
```{r}
names(extractedData)<-gsub("Acc", "Accelerometer", names(extractedData))
names(extractedData)<-gsub("Gyro", "Gyroscope", names(extractedData))
names(extractedData)<-gsub("BodyBody", "Body", names(extractedData))
names(extractedData)<-gsub("Mag", "Magnitude", names(extractedData))
names(extractedData)<-gsub("^t", "Time", names(extractedData))
names(extractedData)<-gsub("^f", "Frequency", names(extractedData))
names(extractedData)<-gsub("tBody", "TimeBody", names(extractedData))
names(extractedData)<-gsub("-mean()", "Mean", names(extractedData), ignore.case = TRUE)
names(extractedData)<-gsub("-std()", "STD", names(extractedData), ignore.case = TRUE)
names(extractedData)<-gsub("-freq()", "Frequency", names(extractedData), ignore.case = TRUE)
names(extractedData)<-gsub("angle", "Angle", names(extractedData))
names(extractedData)<-gsub("gravity", "Gravity", names(extractedData))
names(extractedData)
```
 

5. From the data set in step 4, creates a second, independent tidy data set with the average
of each variable for each activity and each subject.
```{r}
library(data.table)
library(dplyr)
extractedData$Subject <- as.factor(extractedData$Subject)
extractedData <- data.table(extractedData)
tidyData <- aggregate(. ~Subject + Activity, extractedData, mean)
tidyData <- tidyData[order(tidyData$Subject,tidyData$Activity),]
write.table(tidyData, file = "Tidy.txt", row.names = FALSE)
tidyData
```

