---
title: "lab3"
author: "m.pryidun"
date: "2022-11-15"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 3. Tidying data

Data
Dataset hike_data.rds contains information about hiking routes from (Washington Trail
Association)[https://www.wta.org/go-outside/hikes?b_start:int=1].
Data contains the next columns:
name character Name of trail
location character Location of Trail
length character Length of trail (note that most have miles included)
gain character Gain in elevation (Feet above sea level)
highpoint character Highest point in feet above
sea level
rating character User submitted rating (out of 5)
features character Features
description character Description of trail
```{r}
input_data <- readRDS("hike_data.rds")
#input_data
```

Tidying dataset
Create a new dataset clean_hike_trails with the next updates:
1. Convert columns gain, highpoint, rating to numeric values.
```{r}
clean_hike_trails<- transform(input_data, gain = as.numeric(gain), highpoint = as.numeric(highpoint), rating = as.numeric(rating))
```

2. Add new column trip with the type of trip from column length (“roundtrip”, “trails”,
“one-way”).
```{r}
#install.packages("stringr")
library("stringr")
splited_column_lenth<-str_split_fixed(input_data$length, ",",2)
clean_hike_trails$type_of_trip <-splited_column_lenth[, 2]
clean_hike_trails$type_of_trip
```

3. Add new column length_total with the route length from column length, considering
that for “one-way” trip you must double the route length.
```{r}
clean_hike_trails$length_total <- ifelse( 
  splited_column_lenth[, 2]=='one-way',
  splited_column_lenth[, 1]*2,  
  splited_column_lenth[, 1]   
)
clean_hike_trails$length_total
```

4. Add new column location_general with location from column location (a part before
“–”).
```{r}
clean_hike_trails$location_general<-sapply(strsplit(input_data$location, "-"), "[", 1)

clean_hike_trails$location_general
```

5. Add column id with row number
```{r}
clean_hike_trails$ID <- seq.int(nrow(clean_hike_trails))
clean_hike_trails$ID
```


Questioning dataset
Question 1. How many routes have rating more than 4.9
```{r}
length(which(clean_hike_trails$rating >4.9))
```

Question 2. How many routes are “Good for kids” (hint: you can use (unnest function)?

```{r}
v<-unlist(clean_hike_trails$features)
sum(str_detect(v, 'Good for kids'))
```

Question 3. Which unique features can routes have?
```{r}
unique(v)
```

Question 4. What is the most common rating of a route?
```{r}
tail(names(sort(table(clean_hike_trails$rating))), 1)
```

Question 5. Your own question and answer.The highest heightpoint.
```{r}
max(clean_hike_trails$highpoint)
```

